\documentclass[12pt]{report}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{thmtools}
\usepackage[scr]{rsfso}
\usepackage{bm}
% \usepackage{newpxtext,newpxmath}
% \usepackage{mathptmx}
\usepackage[usenames, dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage[
    font=small,
    labelfont=bf,
]{caption}
\usepackage{subcaption}
\usepackage{titlesec}
\usepackage[titles]{tocloft}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage[
    pdftitle={Statistical Depths for Multivariate and Functional Data with Applications},
    pdfauthor={Satvik Saha},
    colorlinks,
    % linkcolor={Blue},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
]{hyperref}
\usepackage[
    hyperref=true,
    style=apa,
    maxcitenames=2,
    url=false,
    eprint=false,
    isbn=false,
    doi=false,
]{biblatex}
\usepackage{fnpct}
\AdaptNote\footcite{oo+m}[footnote]{%
  \setfnpct{dont-mess-around}%
  \IfNoValueTF{#1}
    {#NOTE{#3}}
    {\IfNoValueTF{#2}
       {#NOTE[#1]{#3}}
       {#NOTE[#1][#2]{#3}}}}


\usepackage{lipsum}

\addbibresource{references.bib}

\geometry{a4paper, top = 1in, bottom = 1.5in}
\graphicspath{{images/}}

\titleformat{\chapter}[display]
  {\Huge}
  {\itshape\textcolor{black!70!white}{\chaptertitlename~\thechapter}}
  {0.5em}
  {\scshape\bfseries}

\pagestyle{fancy}
\fancyhead{}
\fancyhead[LO]{\textsc{\nouppercase{\leftmark}}}
\setlength{\headheight}{16pt}

\renewcommand\cftchapfont{\scshape\bfseries}

\newcommand{\chapterstar}[1]{
    \chapter*{#1}
    \addcontentsline{toc}{chapter}{\protect\hphantom{\numberline{\thesection}}#1}
}

\setlength{\parindent}{0em}
\setlength{\parskip}{0.6em}

\input{math}

\title{Statistical Depths for Multivariate and Functional Data with Applications}
\author{Satvik Saha}
\date{}

\begin{document}

    \input{chapters/titlepage}


    \chapter*{Abstract}

    A statistical depth function measures the centrality of a point with
    respect to a distribution, thereby inducing a center-outward ordering on
    the underlying space which in turn generalizes the notions of ranks and
    quantiles.
    This enables the extension of many rank-based nonparametric univariate
    procedures to the multivariate and functional setting.
    In this thesis, we describe the construction of numerous depth functions
    proposed in the literature, both on finite-dimensional as well as function
    spaces, and comment briefly on common properties such as invariance,
    monotonicity, continuity, etc.
    Following this, we explore applications in exploratory data analysis,
    testing, classification, clustering, and outlier detection tasks.
    Finally, we focus on the concept of local depth, and use this to propose a
    new kernel based regression procedure.



    % \chapter*{Dedication}


    \chapter*{Declaration by the Student}

    Date: 20 May 2024\\

    I, \textit{Mr.~Satvik Saha}, Registration No.~\textit{19MS154}, dated
    \textit{20 May 2024}, a student of the \textit{Department of Mathematics
    and Statistics} of the \textit{5 Year BS-MS Dual Degree} programme of
    IISER Kolkata, hereby declare that this thesis is my own work and, to the
    best of my knowledge, it neither contains materials previously published
    or written by any other person, nor has it been submitted for any
    degree/diploma or any other academic award anywhere before.
    I have used the originality checking service to prevent inappropriate
    copying.

    I also declare that all copyrighted material incorporated into this thesis
    is in compliance with the Indian Copyright Act, 1957 (amended in 2012) and
    that I have received written permission from the copyright owners for my
    use of their work.

    I hereby grant permission to IISER Kolkata to store the thesis in a
    database which can be accessed by others.


    \vspace{5em}
    \rule{0.4\textwidth}{0.8pt}

    \vspace{0.1em}
    Satvik Saha

    Department of Mathematics and Statistics,\\
    Indian Institute of Science Education and Research Kolkata.

    Mohanpur 741246, West Bengal, India.

    \chapter*{Certificate from the Supervisor}

    Date: 20 May 2024\\

    This is to certify that the thesis titled \emph{``Statistical Depths for
    Multivariate and Functional Data with Applications''} submitted by
    \emph{Mr.~Satvik Saha}, Registration No.~\emph{19MS154}, dated \emph{20
    May 2024}, a student of the \textit{Department of Mathematics and
    Statistics} of the \textit{5 Year BS-MS Dual Degree} programme of IISER
    Kolkata, is based upon his/her own research work under my supervision.
    I also certify, to the best of my knowledge, that neither the thesis nor
    any part of it has been submitted for any degree/diploma or any other
    academic award anywhere before.
    In my opinion, the thesis fulfils the requirement for the award of the
    degree of \emph{Bachelor of Science and Master of Science}.

    \vspace{5em}
    \rule{0.4\textwidth}{0.8pt}

    \vspace{0.1em}
    Dr.~Anirvan Chakraborty

    Assistant Professor,\\
    Department of Mathematics and Statistics,\\
    Indian Institute of Science Education and Research Kolkata.

    Mohanpur 741246, West Bengal, India.



    \chapter*{Acknowledgments}

    I would like to thank my supervisor Dr.~Anirvan Chakraborty for his
    guidance and support, and for introducing me to this rich topic.
    I also thank my good friend Sohom Gupta, whose discerning eye and
    thoughtful comments have been of great help.


    \tableofcontents


    \chapter{Introduction}
    \label{chap:introduction}
    \input{chapters/introduction}


    \chapter{Multivariate Data}
    \label{chap:multivariate}
    \input{chapters/multivariate}


    \chapter{Functional Data}
    \label{chap:functional}
    \input{chapters/functional}


    \chapter{Local Depth}
    \label{chap:localdepth}
    \input{chapters/localdepth}


    \chapter{Conclusion}
    \input{chapters/conclusion}


    % \nocite{*}
    \printbibliography

    \appendix


\end{document}
