Let $\mathscr{F}$ be a class of distributions on $\R^d$.
It is desirable for a depth function $D\colon \R^d \times \mathscr{F} \to \R$
to satisfy the following properties, described by
\textcite{zuo-serfling-2000}; these are sometimes referred to as the
\emph{Zuo-Serfling axioms}.

\begin{enumerate}
    \item[\textbf{P1}.] \emph{Affine invariance.}
        For any random vector $\vX$ in $\R^d$, any $d \times d$ nonsingular
        matrix $A$, and any $d$-vector $\bm{b}$,
        \begin{equation}
            D(A\vx + \bm{b}, F_{A\vX + \bm{b}}) \;=\; D(\vx, F_{\vX}).
        \end{equation}
        This makes $D(\vx, F_{\vX})$ independent of the choice of coordinate
        system.

    \item[\textbf{P2}.] \emph{Maximality at center.}
        For any $F \in \mathscr{F}$ having `center' $\vth$,
        \begin{equation}
            D(\vth, F) \;=\; \sup_{\vx \in \R^d} D(\vx, F).
        \end{equation}
        This means that the deepest point coincides with some center of
        symmetry of the distribution $F$.

    \item[\textbf{P3}.] \emph{Monotonicity relative to deepest point.}
        For any $F \in \mathscr{F}$ having deepest point $\vth$ and for
        $\alpha \in [0, 1]$,
        \begin{equation}
            D(\vx, F) \;\leq\; D(\vth + \alpha(\vx - \vth), F).
        \end{equation}
        Thus, $D(\Cdot, F)$ monotonically decreases along any ray pointing
        away from the deepest point.

    \item[\textbf{P4}.] \emph{Vanishing at infinity.}
        For any $F \in \mathscr{F}$,
        \begin{equation}
            D(\vx, F) \to 0 \;\text{ as }\;\norm{\vx} \to \infty.
        \end{equation}
\end{enumerate}

By demanding that $D$ be non-negative and bounded, we may assume hereon that
$D$ only takes values in $[0, 1]$.

The notion of a `center' of a distribution in \textbf{P2} is typically
described in terms of symmetry.
We say that a random vector $\vX$ is \emph{centrally symmetric} about
$\vth \in \R^d$ if $\vX - \vth \eqd \vth - \vX$.
Similarly, we say that $\vX$ is \emph{angularly symmetric} about $\vth$ if
$(\vX - \vth) / \norm{\vX - \vth}$ is centrally symmetric about $\bm0$.
An even more restrictive notion of symmetry is \emph{spherical symmetry},
where we demand that $U(\vX - \vth) \eqd \vX - \vth$ for every orthonormal
matrix $U$.
\emph{Elliptical symmetry} requires that $V\vX$ is spherically symmetric about
$\vth$ for some nonsingular matrix $V$.
Finally, the weakest notions of symmetry discussed here is \emph{halfspace
symmetry}, where we impose $P(\vX \in H) \geq 1/2$ for every closed halfspace
in $\R^d$ containing $\vth$.
Thus, the symmetries in decreasing order of strength are $S > E > C > A > H$.

\textcite[Table 2]{mosler-mozharovskyi-2022} provides a detailed summary of
the properties satisfied by the depth functions discussed in the following
section.


\input{chapters/multivariate/depth}

\input{chapters/multivariate/depthcontours}

\input{chapters/multivariate/ddplots}

\input{chapters/multivariate/testing}

\input{chapters/multivariate/classification}

\input{chapters/multivariate/clustering}

% \section{Outlier detection}
